---
title: "logistic_biplot_script"
author: "Camilo Estrella"
date: "2024-02-20"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading libraries

We will start by loading the libraries that will be used, whose function is explained below:

```{r libraries}
library(MultBiplotR)
library(knitr)
library(dplyr)
library(ggplot2)
library(scales)
```

* **MultBiplotR**: It allows multivariate analysis using logistic regression and its graphical representation employing a Biplot.
* **knitr**: The *kable* function is used to obtain a better visualization of the tables.
* **dplyr**: It has multiple functions that allow data management, some of the most interesting are those that allow you to manage tables as if it were a relational database, as is done with SQL.
* **ggplot2**: Perhaps the most important library for the creation of statistical graphics in R.
* **scales**: It is used in conjunction with ggplot2 to scale axes within the plot.

## Loading datasets

The original datasets are 5, named as in the code. In addition, there is a sixth one called "cruzada" (a cross of all 5 tables), which has binary information of aisles of a supermarket in the columns and rows representing the customers. The value 1 means that the customer has shopped in that aisle and 0 the opposite.

The calculation of the "cross" dimension of the dataset is included first.

```{r datasets}
order <- read.csv("Order.csv")
order_product <- read.csv("Order_Product.csv")
product <- read.csv("Product.csv")
aisle <- read.csv("Aisle.csv")
department <- read.csv("Department.csv")

cruzada <- as.data.frame(read.csv("cruzada_df.csv"))
```

```{r cruzada dim}
dimensions <- as.data.frame(matrix(dim(cruzada), ncol = 2))
names(dimensions) <- c("customers (rows)", "aisles (columns)")
kable(dimensions, row.names = FALSE, align = "cc")
```

The datasets are initially from a supermarket and cannot be shared due to source privacy issues.
